Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2018-01-24 manuscript/Appendix4.Rmd

The first seven rows of the fixed effects design matrix (six control plots and one drought plot) is as follows:

```{r fixed_design, echo=F, eval=T, message=F, warning=F}
library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots
source("../code/read_format_data.R") # load data
lmod <- lm(log(anpp) ~ vwc_scaled*Treatment, anpp_data)
x    <- model.matrix(lmod) 
colnames(x) <- c("Int", "VWC", "Drought", "Irrig", "Drought:VWC", "Irrig:VWC")
head(x, 7)
```

@@

\noindent{}The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.

@@

# Figures

```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines. Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots

@@

# References

>>>>

An example of six rows of the fixed effects design matrix (three control plots and three drought plots) is as follows:

```{r fixed_design, echo=F, eval=T, message=F, warning=F}
library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots
source("../code/read_format_data.R") # load data
lmod <- lm(log(anpp) ~ vwc_scaled*Treatment, anpp_data)
x    <- model.matrix(lmod) 
colnames(x) <- c("Int", "VWC", "Drought", "Irrig", "Drought:VWC", "Irrig:VWC")
x[12:17,]
```

@@

\noindent{}The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.
$\Sigma$ is the variance-covariance matrix for the intercept and slope offsets for each plot, which is defined as the among plot variance ($\sigma^2_{plt}$) times the matrix $\textbf{R}$ that defines the correlation between intercept and slope offsets.

@@

```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines. Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE, warning=FALSE, fig.pos="h"}
library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots

@@

### 2018-01-24 manuscript/Appendix5.Rmd

slope_probs <- treat_slopes %>%
  group_by(Treatment, Type) %>%
  summarise(probs = round(get_one_tailed(estimate),2)) %>%
  mutate(prob_text = paste("Pr < 0 =", probs))
slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
slope_probs$xpos <- c(-2.8, -0.6, -2.6, -0.6)
slope_probs$ypos <- c(1.3,1.3,2.55,2.7)

>>>>

slope_probs <- treat_slopes %>%
  group_by(Treatment, Type) %>%
  summarise(probs = round(get_one_tailed(estimate),2)) %>%
  mutate(prob_text = paste("Pr < 0 =", probs))
slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
slope_probs$xpos <- c(-2.8, -0.2, -2.6, -0.2)
slope_probs$ypos <- c(1.3,1.3,2.55,2.7)

### 2018-01-24 manuscript/usses_water_ms.tex

Another problem with relying on historical ecosystem functional
responses to predict impacts of altered precipitation regimes is that
these relationships themselves might shift over the long-term. Shifts in
species identities and/or abundances can alter an ecosystem's functional
response to water availability because different species have different
physiological thresholds. Smith, Knapp, and Collins (2009) introduced
the `Hierarchical Response Framework' for understanding the interplay of
community composition and ecosystem functioning in response to long-term
shifts in resources. In the near term, ecosystem functioning such as
ANPP will reflect the physiological responses of individual species to
the manipulated resource level. For example, ANPP may decline under
simulated drought because the initial community consisted of
drought-intolerant species (Hoover, Knapp, and Smith 2014). Ecosystem
functioning may recover over longer time spans as new species colonize
or initial species reorder in relative abundance. It is also possible
that ecosystem functioning shifts to a new mean state, reflecting the
suite of species in the new community (Knapp, Briggs, and Smith 2012).

Experimental manipulations of limiting resources, like precipitation,
offer a route to understanding how ecosystems will respond to resource
levels that fall outside the historical range of variability (Avolio et
al. 2015; Laureano A. Gherardi and Sala 2015; Knapp et al. 2017).
Altering the amount of precipitation over many years should provide
insight into the time scales at which water-limited ecosystems respond
to chronic resource alteration. We propose four alternative scenarios
for the effect of precipitation manipulation on the ecosystem functional
response to soil moisture based on the Hierarchical Response Framework
(Fig. 1). We define `ecosystem functional response' as the relationship
between available soil moisture and ANPP. The four scenarios are based
on possible outcomes at the community (e.g., community composition) and
ecosystem (e.g., soil moisture-ANPP regression) levels.

First, altered precipitation might have no effect on either ecosystem
functional response or community composition (Fig. 1, top left). In this
case, changes in ANPP would be well predicted by the current, observed
soil moisture-ANPP relationship. This corresponds to the early phases of
the Hierarchical Response Framework, where ecosystem response follows
the physiological responses of individual species. Second, the ecosystem
functional response might change while community composition remains the
same (Fig. 1, top right). A saturating soil moisture-ANPP response fits
this scenario, where individual species hit physiological thresholds or
are limited by some other resource. Third, the ecosystem functional
response might be constant but community composition changes (Fig. 1,
bottom left). In this case, changes in species' identities and/or
abundances occur in response to altered precipitation levels and species
more suited to the new conditions compensate for reduced function of
initial species. Fourth, and last, both ecosystem functional response
and community composition could change (Fig. 1, bottom right). New
species, or newly abundant species, with different physiological
responses completely reshape the ecosystem functional response.

@@

In spring 2011, we established 16 new 1 m\(^2\) plots located in the
\new{largest exclosure at USSES, which also contained} six of our
control plots. We avoided areas on steep hill slopes, areas with greater
than 20\% cover of bare rock, and areas with greater than 10\% cover of
the shrubs \emph{Purshia tridentata} and/or
\emph{Amelanchier utahensis}. We established the new plots in pairs and
randomly assigned each plot in a pair to receive a ``drought'' or
``irrigation'' treatment.
\new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}

@@

\new{We then modeled these daily deviations ($\Delta x_{g,d,\text{trt.}}$)}
from ambient conditions using a linear mixed effects model with
independent variables for treatment (irrigation or drought), season
(winter, spring, summer, fall), rainfall, and all two-way interactions.
Using the local daily weather station data, we recorded rainy days as
any day with measureable precipitation or the day after such a day and
with average temperatures above 3\(^{\circ}\)C (to exclude days with
snowfall). We fit the model using the \texttt{lme4::lmer()} function
(Bates et al. 2015) in R (R Core Team 2016), with random effects for
plot group and date. We weighted observations by the number of unique
sensors or spot measurements that were taken in each plot on that day.
\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{ambient}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{ambient}}$ is the daily ambient soil moisture averaged across all control plots.}
We could only predict soil moisture in the treated plots on days for
which we took at least one ambient soil moisture measurement. This
procedure allowed us to predict daily soil moisture conditions for all
plots, even on those days when some of our direct treatment measurements
were missing due to malfunction of the sensors or dataloggers. See
Kleinhesselink (2017) for more details on our approach to estimating
soil moisture.

@@

We estimated aboveground net primary productivity (ANPP) using a
radiometer to relate ground reflectance to plant biomass (Byrne et al.
2011). We recorded ground reflectance at four wavelengths, two
associated with red reflectance (626 nm and 652 nm) and two associated
with near-infrared reflectance (875 nm and 859 nm). At each plot in each
year, we took four readings of ground reflectances at the above
wavelengths. We also took readings in 12 (2015), 15 (2012, 2013, 2014),
or 16 (2016) calibration plots adjacent to the experimental site, in
which we harvested all aboveground biomass produced in the current year
(we excluded litter and standing dead material), dried it to a constant
weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP. We harvested
at peak green biomass each year, typically in late June.

For each plot and year, we averaged the four readings for each
wavelength and then calculated a greenness index based on the same bands
used to calculate NDVI using the MODIS (Moderate Resolution Imaging
Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer)
bands for NDVI. We regressed the greenness index against the dry biomass
weight from the ten calibration plots to convert the greenness index to
ANPP. We fit regressions to a MODIS-based index and an AVHRR-based index
for each year and retained the regression with the better fit based on
\(R^2\) values. We then predicted ANPP using the best regression
equation for each year (Appendix 3).
\new{Our results do not change when we analyze the soil moisture-NDVI relationship instead of the soil moisture-ANPP relationship (Appendix 5).}

Species composition data came from two sources: yearly census maps for
each plot made using a pantograph (Hill 1920) and yearly counts of
annual species in each plot. From these sources, we determined the
density of all annuals and perennials forbs, the basal cover of
perennial grasses, and the canopy cover of shrubs. We made a large
plot-treatment-year by species matrix, where columns were filled with
either cover or density, depending on the measurement made for the
particular species. We standardized the values in each column so we
could directly compare species quantified with different metrics
(density, basal cover, and canopy cover). This puts all abundance values
on the same scale, meaning that common and rare species are weighted
equally. Assuming that rare species will respond to treatments more than
common ones, our approach is biased towards detecting compositional
changes.
\new{We limited our analysis of community data to observations from the permanent exclosure containing our drought and irrigation treatments (see \textbf{Community composition over time}).
This exclosure included six of the control plots and all of the treatment plots, for a total of 22 plots.
The other eight control plots are in other pastures, which would add a spatial bias to our community analysis focused on temporal dynamics.}

@@

Our main goal was to test whether the relationship between ANPP and soil
moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site
(Blaisdell 1958; Dalgleish et al. 2011; Adler, Dalgleish, and Ellner
2012), we chose to use cumulative volumetric water content from March
through June as our metric of soil moisture (hereafter referred to as
`VWC'). To achieve this goal, we fit a generalized linear mixed effects
regression model with log(ANPP) as the response variable and VWC and
treatment as fixed effects. Plot and year of treatment were included as
random effects to account for non-independence of observations, as
described below. We log-transformed ANPP to reduce heteroscedasticity.
Both log(ANPP) and VWC were standardized to have mean 0 and unit
variance before fitting the model {[}i.e.,
\((x_i - \bar{x})/\sigma_x\){]}.

@@

The results of our five year experiment in a sagebrush steppe
\new{appear to conform to one} of our four predictions. Ecosystem
functional response under chronic \new{drought} was different from the
control treatment, but community composition remained unchanged (Fig.
3A, Fig. 4), representing the top right scenario in Fig. 1.
\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
A strict interpretation of this result implies that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}

\new{However, we do not want to over-interpret the significance of the slope offsets given that the full distributions of}
the slopes of the VWC \(\times\) productivity relationship,
\new{not just the offsets}, were similar among treatments (Table 1). We
therefore conclude that ecosystem functional response is consistent
(\new{similar values}) and weak (\new{all broadly overlapping zero})
across all precipitation treatments.

@@

Second, ANPP at our site may be influenced by additional factors, not
only the cumulative March-June soil moisture covariate we included in
our statistical model. \new{For example,} La Pierre et al. (2016)
\new{found that site-scale ANPP is better predicted by nutrient availability than precipitation.}
Moreover, temperature can impact ANPP directly (Epstein, Lauenroth, and
Burke 1997) and by exacerbating the effects of soil moisture (De Boeck
et al. 2011). Measurements of soil moisture likely contain a signal of
temperature, through its impact on evaporation and infiltration, but the
measurements will not capture the direct effect of temperature on
metabolic and physiological processes. We also did not redistribute snow
across our treatments in the winter, and snow melt may spur early spring
growth. Failure to account for potentially important covariates could
explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the
resulting uncertain relationship we observed between soil moisture and
ANPP across all treatments (Table 1, Fig. A3-2).

Third, the life history traits of the dominant species in our study
ecosystem may explain the consistently positive, but weak and uncertain,
effect of soil moisture on ANPP (Table 1, Fig. 3). Species that live in
variable environments, such as cold deserts, must have strategies to
ensure long-term success as conditions vary. One strategy is bet
hedging, where species forego short-term gains to reduce the variance of
long-term success (Seger 1987). In other words, species follow the same
conservative strategy every year, designed to minimize response to
environmental conditions. The dry and variable environment of the
sagebrush steppe has likely selected for bet hedging species that can
maintain function at low water availability and have weak responses to
high water availability. In so doing, the dominant species in our
ecosystem avoid ``boom and bust'' cycles, which may explain the weak
effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
for the slopes overlapping zero; Table 1).

@@

Our results suggest the species in our focal plant community are
insensitive to to changes in precipitation regime,
\new{at least over the five years of our experiment}. Such insensitivity
can buffer species against precipitation variability in this semi-arid
ecosystem, making them successful in the long run. Longer, chronic
precipitation alteration might reveal plant community shifts that we did
not observe (e.g., Wilcox et al. 2016),
\new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.}
Likewise, a long-term increase in water availability could allow species
that do not bet hedge to gain prominence and dominate the ecosystem
functional response.
\new{The length of the perturbation may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}

>>>>

Another problem with relying on historical ecosystem functional
responses to predict impacts of altered precipitation regimes is that
these relationships themselves might shift over the long-term. Shifts in
species identities and/or abundances can alter an ecosystem's functional
response to water availability because different species have different
physiological thresholds. Smith, Knapp, and Collins (2009) introduced
the `Hierarchical Response Framework' for understanding the interplay of
community composition and ecosystem functioning in response to long-term
shifts in resources. In the near term, ecosystem functioning such as
ANPP will reflect the physiological responses of individual species to
the manipulated resource level. For example, ANPP may initially decline
under simulated drought because the community consisted of
drought-intolerant species (Hoover, Knapp, and Smith 2014), but
functioning may recover over longer time spans as new species colonize
or resident species reorder in relative abundance. It is also possible
that ecosystem functioning eventually shifts to a new mean state,
reflecting the suite of species in the new community (Knapp, Briggs, and
Smith 2012).

Experimental manipulations of limiting resources, like precipitation,
offer a way to test how ecosystems will respond to resource levels that
fall outside the historical range of variability (Avolio et al. 2015;
Laureano A. Gherardi and Sala 2015; Knapp et al. 2017). Altering the
amount of precipitation over many years should provide insight into the
time scales at which water-limited ecosystems respond to chronic
resource alteration. We propose four alternative scenarios for the
effect of precipitation manipulation on the ecosystem functional
response to soil moisture based on the Hierarchical Response Framework
(Fig. 1). We define `ecosystem functional response' as the relationship
between available soil moisture and ANPP. The four scenarios are based
on possible outcomes at the community (e.g., community composition) and
ecosystem (e.g., soil moisture-ANPP regression) levels.

First, altered precipitation might have no effect on either ecosystem
functional response or community composition (Fig. 1, top left). In this
case, changes in ANPP would be well predicted by the current, observed
soil moisture-ANPP relationship. This corresponds to the early phases of
the Hierarchical Response Framework, where ecosystem response follows
the physiological responses of individual species. Second, the ecosystem
functional response might change while community composition remains the
same (Fig. 1, top right). A saturating soil moisture-ANPP response fits
this scenario, where individual species hit physiological thresholds or
are limited by some other resource. Third, the ecosystem functional
response might be constant but community composition changes (Fig. 1,
bottom left). In this case, changes in species' identities and/or
abundances occur in response to altered precipitation levels and species
more suited to the new conditions compensate for reduced function of
initial residents. Fourth, and last, both ecosystem functional response
and community composition could change (Fig. 1, bottom right). New
species, or newly abundant species, with different physiological
responses completely reshape the ecosystem functional response.

@@

In spring 2011, we established 16 new 1 m\(^2\) plots located in the
\new{largest exclosure at USSES, which also contained} six of our
control plots. We avoided areas on steep hill slopes, areas with greater
than 20\% cover of bare rock, and areas with greater than 10\% cover of
the shrubs \emph{Purshia tridentata} and/or
\emph{Amelanchier utahensis}. We established the new plots in pairs and
randomly assigned each plot in a pair to receive the ``drought'' or
``irrigation'' treatment.
\new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}

@@

\new{We then modeled these daily deviations ($\Delta x_{g,d,\text{trt.}}$)}
from ambient conditions using a linear mixed effects model with
independent variables for treatment (irrigation or drought), season
(winter, spring, summer, fall), rainfall, and all two-way interactions.
Using the local daily weather station data, we recorded rainy days as
any day with measureable precipitation or the day after such a day and
with average temperatures above 3\(^{\circ}\)C (to exclude days with
snowfall). We fit the model using the \texttt{lme4::lmer()} function
(Bates et al. 2015) in R (R Core Team 2016), with random effects for
plot group and date. We weighted observations by the number of unique
sensors or spot measurements that were taken in each plot on that day.
\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{ambient}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{ambient}}$ is the daily ambient soil moisture averaged across all control plots.}
We could only predict soil moisture in the treated plots on days for
which we took at least one ambient soil moisture measurement. This
procedure allowed us to predict daily soil moisture conditions for all
plots, even on those days when some of our direct treatment measurements
were missing due to malfunction of the sensors or data loggers. See
Kleinhesselink (2017) for more details on our approach to estimating
soil moisture.

@@

We estimated aboveground net primary productivity (ANPP) using a
radiometer to relate ground reflectance to plant biomass (Byrne et al.
2011). We recorded ground reflectance at four wavelengths, two
associated with red reflectance (626 nm and 652 nm) and two associated
with near-infrared reflectance (875 nm and 859 nm). At each plot in each
year, we took four readings of ground reflectances at the above
wavelengths. We also took readings in 12 (2015), 15 (2012, 2013, 2014),
or 16 (2016) calibration plots adjacent to the experimental site, in
which we harvested all aboveground biomass produced in the current year
(we excluded litter and standing dead material), dried it to a constant
weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP. We made
radiometer measurements and harvested at peak green biomass each year,
typically in late June.

For each plot and year, we averaged the four readings for each
wavelength and then calculated a greenness index based on the same bands
used to calculate NDVI using the MODIS (Moderate Resolution Imaging
Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer)
bands for NDVI. We regressed the greenness index against the dry biomass
weight from the calibration plots to convert the greenness index to
ANPP. We fit regressions to a MODIS-based index and an AVHRR-based index
for each year and retained the regression with the better fit based on
\(R^2\) values. We then predicted ANPP using the best regression
equation for each year (Appendix 3).
\new{Our results do not change when we analyze the soil moisture-NDVI relationship instead of the soil moisture-ANPP relationship (Appendix 5).}

Species composition data came from two sources: yearly census maps for
each plot made using a pantograph (Hill 1920) and yearly counts of
annual species in each plot. From these sources, we determined the
density of all annuals and perennials forbs, the basal cover of
perennial grasses, and the canopy cover of shrubs. We made a large
plot-treatment-year by species matrix, where columns were filled with
either cover or density, depending on the measurement made for the
particular species. We standardized the values in each column so we
could directly compare species quantified with different metrics
(density, basal cover, and canopy cover). This puts all abundance values
on the same scale, meaning that common and rare species are weighted
equally. Assuming that rare species will respond to treatments more than
common ones, our approach is biased towards detecting compositional
changes.
\new{We limited our analysis of community data to observations from the permanent exclosure containing our drought and irrigation treatments (see \textbf{Community composition over time}).
This exclosure included six of the control plots and all of the treatment plots, for a total of 22 plots.
The other eight control plots are in other pastures. 
Including these plots would add spatial variation in composition, complicating our goal of describing temporal trends in composition.}

@@

Our main goal was to test whether the relationship between ANPP and soil
moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site
(Blaisdell 1958; Dalgleish et al. 2011; Adler, Dalgleish, and Ellner
2012), we chose to use cumulative volumetric water content from March
through June as our metric of soil moisture (hereafter referred to as
`VWC'). We fit a generalized linear mixed effects regression model with
log(ANPP) as the response variable and VWC and treatment as fixed
effects. Plot and year of treatment were included as random effects to
account for non-independence of observations, as described below. We
log-transformed ANPP to reduce heteroscedasticity. Both log(ANPP) and
VWC were standardized to have mean 0 and unit variance before fitting
the model {[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.

@@

We found that ecosystem functional response under chronic \new{drought}
was different from the control treatment, but community composition
remained unchanged (Fig. 3A, Fig. 4), representing the top right
scenario in Fig. 1.
\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
A strict interpretation of this result implies that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}

\new{However, we do not want to over-interpret the significance of the slope offsets given that the overall slopes}
of the VWC \(\times\) productivity relationship,
\new{not just the offsets}, were similar among treatments (Table 1). We
therefore conclude that ecosystem functional response is consistent
(\new{similar values}) and weak (\new{all broadly overlapping zero})
across all precipitation treatments.

@@

Second, ANPP at our site is likely influenced by additional factors, not
only the cumulative March-June soil moisture covariate we included in
our statistical model. \new{For example,} La Pierre et al. (2016)
\new{found that site-scale ANPP is better predicted by nutrient availability than precipitation.}
Moreover, temperature can impact ANPP directly (Epstein, Lauenroth, and
Burke 1997) and by exacerbating the effects of soil moisture (De Boeck
et al. 2011). Measurements of soil moisture likely contain a signal of
temperature, through its impact on evaporation and infiltration, but the
measurements will not capture the direct effect of temperature on
metabolic and physiological processes. We also did not redistribute snow
across our treatments in the winter, and snow melt may spur early spring
growth. Failure to account for potentially important covariates could
explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the
resulting uncertain relationship we observed between soil moisture and
ANPP across all treatments (Table 1, Fig. A3-2).

Third, the life history traits of the dominant species in our study
ecosystem may explain the consistently positive, but weak and uncertain,
effect of soil moisture on ANPP (Table 1, Fig. 3). Species that live in
variable environments, such as cold deserts, must have strategies to
ensure long-term success as conditions vary. One strategy is bet
hedging, where species forego short-term gains to reduce the variance of
long-term success (Seger 1987). In other words, species follow the same
conservative strategy every year, designed to minimize losses during
unfavorable periods. The dry and variable environment of the sagebrush
steppe has likely selected for bet hedging species that can maintain
function at low water availability and have weak responses to high water
availability. In so doing, the dominant species in our ecosystem avoid
``boom and bust'' cycles, which may explain the weak effect of soil
moisture on ANPP (i.e., the Bayesian credible intervals for the slopes
overlapping zero; Table 1).

@@

Our results suggest that the species in our focal plant community are
insensitive to to changes in precipitation regime,
\new{at least over the five years of our experiment}. Such insensitivity
can buffer species against precipitation variability in this semi-arid
ecosystem, making them successful in the long run. Longer, chronic
precipitation alteration might reveal plant community shifts that we did
not observe (e.g., Wilcox et al. 2016),
\new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.}
Likewise, a long-term increase in water availability could allow species
that do not bet hedge to gain prominence and dominate the ecosystem
functional response.
\new{The length of the perturbation may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
